Choosing a Population Genomics Approach

### 2021-12-17 choosing-population-genetics.Rmd
Choosing a popgen approach corrections

# Population genomic sequencing approaches
Deciding on a sequencing approach for a new project can be challenging, and depends primarily on the questions you are interested in, but also constraints set by funding, existing genomic resources, sample size, genome characteristics, and expertise in the research group. The main categories of population genomic sequencing approaches are:

### 1) High coverage individual whole genome resequencing (hcWGS)  {.unlisted .unnumbered}
**Whole-genome resequencing (WGS)** is where whole genomes of multiple individuals are sequenced and mapped to an assembled reference genome in order to compare variable sites between individuals or populations. **Coverage depth** refers to the average number of sequencing reads per region of the genome. With more sequencing reads at a region (higher coverage depth), errors that arise during sequencing have a smaller impact on accurately identifying variable sites (termed "genotyping") [@Fuentes-Pardo2017-nu].

### 2) Low coverage individual WGS (lcWGS)  {.unlisted .unnumbered}
Low coverage whole genome sequencing is where whole genomes of many individuals are sequenced, but depth of coverage so low that individual genotypes cannot be confidently assigned (cite). Instead of assigning specific, "hard called genotypes, lcWGS data requires probabilistic analyses that account for uncertainty about the true genotype and can incororate uncertanty accross all sequenced samples. Typically, sequencing depth of > 3 reads is considered "low coverage".

### 3) Reduced-representation sequencing {.unlisted .unnumbered}
As the name suggests, this approach aims to sequence a reduced portion of the genome across multiple individuals at moderate to high coverage for variant discovery and genotyping with high confidence. The genome fragments can be selected at random, or through targeted probe assays.
#### RADseq/Genotype-by-sequencing  {.unlisted .unnumbered}
This very popular class of methods involve sequencing a subset of genomic regions randomly distributed throughout the genome, often using restriction enzymes (i.e., RADseq: restriction site-associated DNA). 
#### Targeted capture methods (e.g., exome capture)  {.unlisted .unnumbered}

### 4) Pooled individuals (Poolseq)  {.unlisted .unnumbered}

@@

It should be noted that you don't need to choose only one approach for all samples! If you are starting a research program on a species without any genomic resources, you may want to start with de novo RADseq to understand basic population structure and genetic diversity, then once a reference genome is assembled move to high/low coverage WGS. In the [MarineOmics seminar](link), Misha Matz recommended choosing one individual to use for genome assembly, doing high coverage WGS on a few other individuals that represent most of the variation in your taxa, and then WGS many other individuals at low coverage. The high coverage individuals can be useful for imputing missing genotypes in the low coverage samples.

This table briefly summarizes the pros and cons of five sequencing approaches and their appropriateness for answering specific questions. If you are interested in answering multiple questions (eg, neutral population structure and adaptive variation) then you would often want a method that can do both (eg, WGS).

@@

# Recommended readings  
Below we list a number of great review articles and empirical comparison studies that can help when making the decision. This topic was also discussed during the MarineOmics seminars (link).

@@

* [@Benjelloun2019-dq]: empirical comparison of high coverage WGS, low coverage WGS, random variants (eg. from RADseq), exome capture, and commercial SNP chip for mammals (genome size=2.6Gb). For neutral genomic diversity, found 5k-10k random variants were enough to accurately estimate, but comercial panales/exome capture had acertainment bias. To detect selection and accurate estimates of LD, at least 1M variants were required. For the studied species, 5x coverage was just as accurate as 12x coverage. Low coverage WGS (< 5x) should use genotype likelihood methods, however the rate of false positives for heterozygotes is higher than classic genotyping methods at higher coverage.  
* [@Dorant2019-fb]: Case study comparing Poolseq vs. reduced representation seq (GBS) vs Rapture (targeted capture of RAD loci) for inferring population structure with weak genetic differentiation. All three detected same weak pop structure pattern, but Poolseq gave Fst values 3-5x higher than Rapture/GBS.

@@

# References

>>>>

# Population genomic sequencing approaches
Deciding on a sequencing approach for a new project can be challenging. Your choice depends primarily on the questions of interest, but also constraints set by funding, existing genomic resources, sample size, genome characteristics and expertise in the research group. The main categories of population genomic sequencing approaches include (1) High coverage individual whole-genome resequencing (hcWGS) and (2) Low coverage individual WGS (lcWGS), (3).

**Coverage depth** refers to the average number of sequencing reads per region of the genome (distinct from coverage breadth, see [@sims2014sequencing]). With more sequencing reads at a region (higher coverage), errors that arise during sequencing have a smaller impact on accurately identifying variable sites (termed “genotyping”) [@Fuentes-Pardo2017-nu].

### 1) High coverage individual whole-genome resequencing (hcWGS)  {.unlisted .unnumbered}
**Whole-genome resequencing (WGS)** is where whole genomes of multiple individuals are sequenced and mapped to an assembled reference genome in order to compare variable sites between individuals or populations. This approach is considered the is the 'gold standard' for DNA resequencing because it includes all variant types (i.evSNVs, indels, structural variants and CNVs) [@sims2014sequencing].

### 2) Low coverage individual WGS (lcWGS)  {.unlisted .unnumbered}
Low coverage whole genome sequencing is where whole genomes of many individuals are sequenced, but depth of coverage is so low (sequencing depth < 3) that individual genotypes cannot be confidently assigned [@korneliussen2014angsd]. Instead of assigning specific, “hard called" genotypes, lcWGS data requires probabilistic analyses that account for uncertainty about the true genotype and can incorporate uncertainty across all sequenced samples.

### 3) Reduced-representation sequencing {.unlisted .unnumbered}
As the name suggests, this approach aims to sequence a reduced portion of the genome across multiple individuals at moderate to high coverage for variant discovery and genotyping with high confidence. The genome fragments can be selected at random, or through targeted probe assays. Variations of this type of sequencing approach include:

#### RADseq/Genotype-by-sequencing  {.unlisted .unnumbered}
A class of methods involving sequencing a subset of genomic regions randomly distributed throughout the genome, often using restriction enzymes (i.e. restriction site-associated DNA (RADseq)).

#### Targeted capture methods (e.g., whole-exome sequencing (WES))  {.unlisted .unnumbered}
Sequence capture methods use a set of probes designed with a-priori knowledge to focus sequencing effort on a set of hundreds to tens of thousands of specific loci. Several capture approaches for non-model species can utilise other sequencing strategies to design probes corresponding to genomic regions (eg. transcriptome sequencing, ddRAD) (see [@jones2016targeted] for a review).

### 4) Pooled individuals (Poolseq)  {.unlisted .unnumbered}
Sequencing pools of individuals to provide information on genome wide  population  allele  frequencies (see review in [@futschik2010next]).

@@

It should be noted that you don’t need to choose only one approach for all samples! If you are starting a research program on a species without any genomic resources, you may want to start with denovo RADseq to understand basic population structure and genetic diversity, then once a reference genome is assembled move to high/low coverage WGS.In the [MarineOmics seminar](https://marineomics.github.io/panels.html), Misha Matz recommended choosing one individual to use for genome assembly, doing high coverage WGS on a few other individuals that represent most of the variation in your taxa, and then WGS many other individuals at low coverage. The high coverage individuals can be useful for imputing missing genotypes in the low coverage samples.

The table below briefly summarizes the pros and cons of five sequencing approaches and their appropriateness for answering specific questions. If you are interested in answering multiple questions (i.e. neutral population structure and adaptive variation) then you would often want a method that can do both (i.e. WGS).

@@

# Recommended readings  
Below we list a number of great review articles and empirical comparison studies that can help when making the decision. This topic was also discussed during the [MarineOmics seminars](https://marineomics.github.io/panels.html).

@@

* [@Benjelloun2019-dq]:an empirical comparison of high coverage WGS, low coverage WGS, random variants (eg. from RADseq), exome capture, and commercial SNP chip for mammals (genome size=2.6Gb). For neutral genomic diversity, found 5k-10k random variants were enough to accurately estimate, but commercial panels/exome capture had ascertainment bias. To detect selection and accurate estimates of LD, at least 1M variants were required. For the studied species, 5x coverage was just as accurate as 12x coverage. Low coverage WGS (< 5x) should use genotype likelihood methods, however, the rate of false positives for heterozygotes is higher than classic genotyping methods at higher coverage.
* [@Dorant2019-fb]:  A case study comparing Poolseq vs. reduced representation seq (GBS) vs Rapture (targeted capture of RAD loci) for inferring population structure with weak genetic differentiation. All three detected the same weak pop structure pattern, but Poolseq gave Fst values 3-5x higher than Rapture/GBS.

@@
